Fitness tracking reveals task-specific associations between memory, mental health, and exercise

### 2021-10-23 paper/main.tex
updated author affiliations
typo fixes

\title{Fitness tracking reveals task-specific associations between
  memory, mental health, and exercise}
\author[1, $\star$]{Jeremy R. Manning}
\author[1,2]{Gina M. Notaro}
\author[1]{Esme Chen}
\author[1]{Paxton C. Fitzpatrick}
\affil[1]{Dartmouth College, Hanover, NH}
\affil[2]{Lockheed Martin, Bethesda, MD}
\affil[$\star$]{Address correspondence to
  jeremy.r.manning@dartmouth.edu}

@@

\begin{abstract}
  Physical exercise can benefit both physical and mental well-being.
  Different forms of exercise (i.e., aerobic versus anaerobic; running
  versus walking, swimming, or yoga; high-intensity interval training
  versus endurance workouts; etc.) impact physical fitness in
  different ways.  For example, running may substantially impact leg
  and heart strength but only moderately impact arm strength. We
  hypothesized that the mental benefits of exercise might be similarly
  differentiated.  We focused specifically on how different forms of
  exercise might related to different aspects of memory and mental
  health.  To test our hypothesis, we collected (in aggregate) roughly
  a century's worth of fitness data.  We then asked participants to
  fill out surveys asking them to self-report on different aspects of
  their mental health.  We also asked participants to engage in a
  battery of memory tasks that tested their short and long term
  episodic, semantic, and spatial memory performance.  We found that
  participants with similar exercise habits and fitness profiles
  tended to also exhibit similar mental health and task performance
  profiles.  These effects were task-specific in that different
  exercise patterns or fitness characteristics varied with different
  aspects of memory, on different tasks.
\end{abstract}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/weighted_timecourse_summary}
\caption{\textbf{Dynamics of physical activity varies with memory
    performance and mental health measures.}  \textbf{a. Daily step counts.}  Each
  timecourse is weighted by either performance on immediate recall tests
  (left panel) or on delayed recall tests (right panel).  The black
  (baseline) timecourses display the (unweighted) average across all
  participants.  \textbf{b. Daily duration (in minutes) of low-intensity physical
    activity.}  Timecourses are displayed in the same format and color
scheme as those in Panel A.  Analogous timecourses for additional
fitness-related measures may be found in Figures~\activityTimecourse,
\cardioTimecourse, and \sleepTimecourse.  \textbf{c. Timecourses of physical
  activity, weighted by mental health measures.}  The timecourses in
each panel display the average daily step counts (top panel) or
duration of low-intensity activity (bottom panel).  The colored lines
show average activity dynamics weighted by self-reported stress levels
at the start of the experiment (purple) and self-reported ``typical''
stress levels (pink).  The baseline curves (black) display the
average across all participants (re-plotted in Panel C to illustrate
scale differences across panels).  Timecourses for additional mental
health-related and fitness-related measures may be found in
Figures~\activityTimecourseMH, \cardioTimecourseMH, and \sleepTimecourseMH. Error ribbons in all panels denote
the standard error of the mean.  Horizontal lines below each panel's
timecourses denote intervals over which each weighted measure (color) differs
from the unweighted baseline (via a paired sample two-sided $t$-test
of the weighted mean values for each measure within a 30 day window around each timepoint; horizontal lines denote $p <
0.05$, corrected.}
\label{fig:dynamics}
\end{figure}

>>>>

\title{Fitness tracking reveals task-specific associations between
  memory, mental health, and exercise}
\author[1, $\star$]{Jeremy R. Manning}
\author[1]{Gina M. Notaro}
\author[1]{Esme Chen}
\author[1]{Paxton C. Fitzpatrick}
\affil[1]{Dartmouth College, Hanover, NH}
\affil[$\star$]{Address correspondence to
  jeremy.r.manning@dartmouth.edu}

@@

\begin{abstract}
  Physical exercise can benefit both physical and mental well-being.
  Different forms of exercise (i.e., aerobic versus anaerobic; running
  versus walking, swimming, or yoga; high-intensity interval training
  versus endurance workouts; etc.) impact physical fitness in
  different ways.  For example, running may substantially impact leg
  and heart strength but only moderately impact arm strength. We
  hypothesized that the mental benefits of exercise might be similarly
  differentiated.  We focused specifically on how different forms of
  exercise might relate to different aspects of memory and mental
  health.  To test our hypothesis, we collected (in aggregate) roughly
  a century's worth of fitness data.  We then asked participants to
  fill out surveys asking them to self-report on different aspects of
  their mental health.  We also asked participants to engage in a
  battery of memory tasks that tested their short and long term
  episodic, semantic, and spatial memory performance.  We found that
  participants with similar exercise habits and fitness profiles
  tended to also exhibit similar mental health and task performance
  profiles.  These effects were task-specific in that different
  exercise patterns or fitness characteristics varied with different
  aspects of memory, on different tasks.
\end{abstract}

@@

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/weighted_timecourse_summary}
\caption{\textbf{Dynamics of physical activity varies with memory
    performance and mental health measures.}  \textbf{a. Daily step counts.}  Each
  timecourse is weighted by either performance on immediate recall tests
  (left panel) or on delayed recall tests (right panel).  The black
  (baseline) timecourses display the (unweighted) average across all
  participants.  \textbf{b. Daily duration (in minutes) of low-intensity physical
    activity.}  Timecourses are displayed in the same format and color
scheme as those in Panel A.  Analogous timecourses for additional
fitness-related measures may be found in Figures~\activityTimecourse,
\cardioTimecourse, and \sleepTimecourse.  \textbf{c. Timecourses of physical
  activity, weighted by mental health measures.}  The timecourses in
each panel display the average daily step counts (top panel) or
duration of low-intensity activity (bottom panel).  The colored lines
show average activity dynamics weighted by self-reported stress levels
at the start of the experiment (purple) and self-reported ``typical''
stress levels (pink).  The baseline curves (black) display the
average across all participants (re-plotted in Panel C to illustrate
scale differences across panels).  Timecourses for additional mental
health-related and fitness-related measures may be found in
Figures~\activityTimecourseMH, \cardioTimecourseMH, and \sleepTimecourseMH. Error ribbons in all panels denote
the standard error of the mean.  Horizontal lines below each panel's
timecourses denote intervals over which each weighted measure (color) differs
from the unweighted baseline (via a paired sample two-sided $t$-test
of the weighted mean values for each measure within a 30 day window around each timepoint; horizontal lines denote $p <
0.05$, corrected).}
\label{fig:dynamics}
\end{figure}
